### 2016-07-12 cross_compilation.tex
Separate automation tools from cross-compiling

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary task execution in the \ac{Raspi}s. Given the
computing capabilities of the \ac{Raspi}, common computational demanding tasks
such as source code compilation and library linking can be challenginng if
not prohibitive in terms of \ac{CPU} or memory usage. Thus, we also include
a description of how to cross-compile source code from the testbed
administrator \ac{PC} to the \ac{Raspi}s. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally in a \ac{Raspi} making
necessary to use a keyboard and a monitor. To do so, the testbed
administrator needs to put the memory cards in the \ac{Raspi}s and turn them
on, for them to be able to boot. The devices should boot properly and have
connectivity if the image and \ac{DHCP} server were properly set previously.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to many repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to and / or from multiple devices, etc. To
simplify the management and avoid execution issues, Fabric provides
a Python library to ease the major task of working with multiple devices from
a single \ac{PC}. First, the testbed administrator creates a directory to
hold the \texttt{Fabric} source code:

\begin{lstlisting}[]
$ export CODEDIR="${HOME}/code"
$ mkdir -p ${CODEDIR}
$ cd ${CODEDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Then, the \texttt{\$CODEDIR/fabfile.py} file below provides a script with
some basic functionalities that can perform the few items above (i-iii).
In general, other administrators may require different functionalities
but that goes out of the scope of this work. Nevertheless, the following
file serves as a starting boilerplate:

% Exit chroot, umount, and write to memory card
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$CODEDIR/fabfile.py}@>
<@\textcolor{gray}{
-------------------------------------}
\Reactivatenumber @>
from fabric.api import env, task, sudo
# Python Fabric script to run commands on multiple hosts through ssh
#
# Run script as 'fab <task>', where <task> is one of the scripts functions
# marked as a tesk. The task marked as 'default' will be run if <task> is not
# specified

env.hosts = ['rasp00.domain.com','rasp01.domain.com','rasp02.domain.com']
env.user = 'pi'
env.password = 'raspberry'

@task
def reboot():
    """ Reboot device """
    sudo('reboot', quiet=True)

@task
def install(program):
    """
    Install a program
    program: program name
    """
    result = sudo('apt-get install -y {}'.format(program), quiet=True)
    print(result)

@task
def push(src,dst):
    """
    Copy file to device
    src: source file path
    dst: destination file path
    """
    put(src, dst)

\end{lstlisting}
\FloatBarrier

The previous \texttt{fabfile} shows three function that perform our
example tasks. These functions utilize variables and subsequent functions
from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
among others. Each of these \ac{API} functions permit to define environment
variables, create the administrator tasks through decorators or run the
mentioned task in \texttt{sudo} mode, respectively.
When a task is called from the terminal, Fabric searches the directory for
the \texttt{fabfile.py} file and executes the desired task. The syntax for
executing a task with arguments is in the form
\texttt{fab <FUNCTION>:arg1,arg2,...}. Some executions from the terminal
are shown as follows \todot{Test the following commands later and print the
outputs}:

\begin{lstlisting}[]
$ fab reboot
$ fab install:tmux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The first function above reboots the \ac{Raspi}s in the lists of hosts and
the second function installs a program given by an argument. In both cases,
after a successful \ac{SSH} connection is made, Fabric employs the
\ac{Raspi}'s \texttt{reboot} and \texttt{apt-get} commands in \texttt{sudo}
mode to do the required tasks. Below it is shown an example for
the \texttt{push} task which uses two arguments. Here we copy
\texttt{my\_file} from the testbed administrator \ac{PC} to all the host
\ac{Raspi}s:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/my_file",'~/'
[192.168.87.107] Executing task 'push'
[192.168.87.107] put: /home/<USER>/code/my_file -> /home/pi/my_file

Done.
Disconnecting from 192.168.87.107... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Fabric has many other functionalities that are useful in
controlling a large set of \ac{Raspi}s. For example, we may extract files
or run automated experiments. The included functionalities in
the \texttt{\$CODEDIR/fabfile.py} file will depend on the requirements
of the testbed administrator.

%execute a script function by calling "fab install:feh"
\subsection{Long running jobs using SSH}
There are times when a task may need to run for several hours or even days
in the \ac{Raspi}s, particularly when related to simulations or measurement
campaigns. For this purpose, it might be necessary to keep open a \ac{SSH}
connection in the \ac{Raspi}s without risking that the connection will
be interrupted and a given \ac{Raspi} terminating the task.

There are methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly. One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a shell
window, a \texttt{screen} session, which does not terminate even with
connectivity interruptions. Users can attach and detach from such a session
as desired. The following procedure presents how to use \texttt{screen} with
\ac{SSH} to: (i) Login to a \ac{Raspi}, (ii) open a screen session, (iii)
execute an example command, (iv) detach from the screen session, (v)
terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again
and (vii) attach to screen session to see the program still running.
From the testbed administrator, we start by establishing \ac{SSH} connection
to a \ac{Raspi} and open a screen session:
\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
$ screen
Screen version 4.02.01 (GNU) 28-Apr-14
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
Press space or return to clear the shell.

At this stage, we are in a screen session. Now execute a program that never ends:
\begin{lstlisting}[]
$ top
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

When \texttt{top} is running, press Ctrl+a Crtl+d to detach from
the \texttt{screen} session.

Terminate the SSH connection and login again:
\begin{lstlisting}[]
$ exit
$ SSH pi@<RASP_IP>
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Now that we are logged in to the \ac{Raspi} again, lets see if a
\texttt{screen} session is still running:
\begin{lstlisting}[]
$ screen -list
There is a screen on:
  1668.pts-0.rasp09 (06/06/16 04:02:15) (Detached)
1 Socket in /var/run/screen/S-pi.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

It is seen that the session is still running and that no user is currently
attached to the session. Attach to the session:
\begin{lstlisting}[]
$ screen -r 1668.pts-0.rasp09
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

After attaching again, \texttt{top} should still be running. \texttt{screen}
has a few more functionalities that is outside the scope of this
paper.

To terminate the \texttt{screen} session. First, terminate \texttt{top}
by pressing 'q', and then type 'exit' two times to first exit the
\texttt{screen} session and then terminate the \ac{SSH} connection:

%Problem when you logout, then applications will terminate. SCREEN is the answer.
\subsection{Cross-compilation: From your PC to the Raspberry Pi}
Some processing expensive tasks such as compiling code libraries can with
advantage be performed in a \ac{PC} rather than a \ac{Raspi} to save time.
This section will present how to setup a toolchain in a \ac{PC} that may
be used to cross compile C++ code to run in \ac{Raspi}s.
There will be given an example of compiling a simple C++ application to run
on \ac{Raspi}, but also how Kodo-cpp can be cross compiled for \ac{Raspi}.

%Some tasks can with advantage be performed in \ac{PC} rather than a \ac{Raspi}
%due to the
%
%Modern \ac{PC}s are significantly better equipped than the \ac{Raspi}s that in
%contrasts to \ac{PC}s target a far more affortable and smaller module size. This
%means that it may be preferred or even mandatory to perform some computational
%expensive tasks in a \ac{PC} rather than in \ac{Raspi}s.
%Example of tasks that are quite computational expensive is to compile software
%packages and large libraries.
%Another reason to use a \ac{PC} may be the hardware limitations of \ac{Raspi}s
%in terms of e.g. memory and disk space. Some software packages might simply be
%too memory demanding to be compiled in a \ac{Raspi}.
%Thus, this section will present how to setup a toolchain in a \ac{PC} that can
%be used to cross compile C++ code to the ARM architecture that the \ac{Raspi}s.
%The toolchain is mandatory on most consumer \ac{PC}s due to the different
%processor architecture.
%
%Furthermore, we will cover how to cross compile a simple C++ example and
%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
%\ac{Raspi} and executed.
%There exists bindings to Kodo in various programming language. We will present
%how to also run a Kodo-python application in a \ac{Raspi}.

%WRITE INTRO OF WHY WE WANT TO CROSS COMPILE. E.G. 1) WE DO NOT HAVE MONITORS
%CONNECTED TO THE RASPIS, 2) A PC IS SIGNIFICANTLY FASTER AND FAR BETTER EQUIPPED (RAM, CPU, HARDDRIVE), 3) MORE CONVINIENT TO DO EVERYTHING IN ONE PC AND DISTRIBUTE IT TO THE RASPIS AFTERWARDS. MORE?

\subsubsection{Setup toolchain on PC}

@@

\subsubsection{Cross compile example}

@@

\subsubsection{Cross compile Kodo}

>>>>

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and / or compilation time. Thus, in
this Section we present a procedure of how to cross-compile C++
source code from the testbed administrator \ac{PC} for the \ac{ARM}
architecture of the \ac{Raspi}s. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator \ac{PC} in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program to run
on a \ac{Raspi}.

Furthermore, given that our testbed purpose is for
network coding applications, we also present how to cross-compile Kodo
\cite{kodo2011pedersen}, a C++11 network coding library to perform
encoding, decoding and recoding operations. In this way, we aim to present
a fully configurable and manageable testbed with the capabilities to
evaluate network coding protocols with several \ac{Raspi}s and locally
store measurements from different evaluations. Therefore, we also show
how \texttt{kodo-cpp}, a set of high-level C++ bindings for Kodo, can be
cross-compiled for applications with the \ac{Raspi}.

%The toolchain is mandatory on most consumer \ac{PC}s due to the different
%processor architecture.
%
%Furthermore, we will cover how to cross compile a simple C++ example and
%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
%\ac{Raspi} and executed.

\subsection{Setup toolchain on PC}

@@

\subsection{Cross compile example}

@@

\subsection{Cross compile Kodo}
